### 2012-01-22 notes.tex
Update paper with description of statistics.  Add TODO.

Define the event history $\mathcal{A}_t = \{e_m: t_m \in [0,t)\}$.
For a given dyad $(i,j)$ and time $t$ we may also have a vector of statistics $s_{ij}(t|\mathcal{A}_t,\mathbf{z})$ that is a function of the previous history of events $\mathcal{A}_t$ and the latent cluster assignments $z$.  Note the statistics $s_{ij}$ also do not change until an event involving either $z_i$ or $z_j$ occurs.  We model the intensity functions via a log linear model conditioned on this information
\begin{align}
\log \lambda_{ij}(t | \mathcal{A}_t,\mathbf{z}) = \boldsymbol{\beta}_{z_i,z_j} \mathbf{s}_{ij}(t|\mathcal{A}_t,\mathbf{z}).
\end{align}
where for each pair of clusters $(z_i,z_j)$ we have a vector of parameters $\beta_{z_i,z_j}$ that corresponds to the covariate vector $s_{ij}(t|\cdot)$.  This is reminiscent of the stochastic blockmodel \cite{Nowicki2001, Kemp2006} which models the probability of a dyad as $p(y_{ij}) =\mbox{logit}^{-1}( \eta_{z_i,z_j})$ where $\eta_{z_i,z_j}$ is essentially a mixing rate between group $z_i$ and group $z_j$.  Here, however, we use a blockmodel structure facilitates the study of intra-group and inter-group dynamics via a continuous-time network model.

@@

We include a similar set of statistics in $s_{ij}$ as used in \cite{Butts2008,Vuy2011}.
Note that the model assumptions imply $s_{ij}(t|\mathcal{A}_t,\mathbf{z}) = f(\mathcal{A}$.

Let $r$ be the last event from a $z_i$ node to a $z_j$ node, that is $r = \mbox{argmax}_{r \in \{1, \ldots, m\}}I(t_r < t, z_{i_r} = z_i, z_{j_r} = z_j)$

\begin{itemize}
\item In and out degree of sender and receiver
\item Reciprocity
\item Transitivity, triangle closure
\item Shared contacters, contactees
\end{itemize}

TODO: Would like things that combine shared contacts with recency. (i.e. Facebook-like terms.)

@@

\begin{enumerate}
\item $t_0 = 0$, all $s_{ij} = 1$
\item For $i = 1, \ldots, N$, draw $z_i \sim \mbox{Categorical}(\theta)$
\item For $m = 1, \ldots, M$:
  \begin{enumerate}
  \item For all $(i,j)$ such that $z_i,z_j \in \{z_{i_{m-1}},z_{j_{m-1}}\}$:
    \begin{enumerate}
    \item Compute $s_{ij}^*$ using events $1$ through $m-1$
    \item Set $\lambda_{ij}^* \leftarrow \exp\{ \beta_{z_i,z_j} s_{ij}^*\}$ 
    \end{enumerate}
  \item Draw $t_m \sim \mbox{Exponential}\left(\sum_{ij} \lambda_{ij}^*\right)$
  \item Draw $(i_m,j_m) \sim \mbox{Categorical}\left(\lambda_{ij}^* / \sum_{ij}\lambda_{ij}^*\right)$
  \item Save $s_{ij}(t_m) \leftarrow s_{ij}^*$
  \item Save $\lambda_{ij}(t_m) \leftarrow \lambda_{ij}^*$
  \end{enumerate}
\end{enumerate}

>>>>

Define the event history $\mathcal{A}_t = \{e_m: t_m \in [0,t)\}$.
For a given dyad $(i,j)$ and time $t$ we may also have a vector of statistics $s_{ij}(t|\mathcal{A}_t,\mathbf{z})$ that is a function of the previous history of events $\mathcal{A}_t$ and the latent cluster assignments $z$.  Note the statistics $s_{ij}$ also do not change until an event involving either $z_i$ or $z_j$ occurs.  We model the intensity functions via a log linear model conditioned on this information
\begin{align}
\log \lambda_{ij}(t | \mathcal{A}_t,\mathbf{z}) = \boldsymbol{\beta}_{z_i,z_j} \mathbf{s}_{ij}(t|\mathcal{A}_t,\mathbf{z}).
\end{align}
where for each pair of clusters $(z_i,z_j)$ we have a vector of parameters $\beta_{z_i,z_j}$ that corresponds to the covariate vector $s_{ij}(t|\cdot)$.  This is reminiscent of the stochastic blockmodel \cite{Nowicki2001, Kemp2006} which models the probability of a dyad as $p(y_{ij}) =\mbox{logit}^{-1}( \eta_{z_i,z_j})$ where $\eta_{z_i,z_j}$ is essentially a mixing rate between group $z_i$ and group $z_j$.  Here, however, the blockmodel structure facilitates the study of intra-group and inter-group dynamics via a continuous-time network model.

@@

We include a similar set of statistics in $s_{ij}$ as used in \cite{Butts2008,Vuy2011}.  Let $v_{ijm}$ be the index of the last event where $(i,j)$ occurred; if $(i,j)$ has never occurred, we set $v_{ijm}=0$.
%Note that the model assumptions imply $s_{ij}(t|\mathcal{A}_t,\mathbf{z}) = f(\mathcal{A}$.

%Let $r$ be the last event from a $z_i$ node to a $z_j$ node, that is $r = \mbox{argmax}_{r \in \{1, \ldots, m\}}I(t_r < t, z_{i_r} = z_i, z_{j_r} = z_j)$

\begin{enumerate}
\item Sender out-degree: $s_{ij1}(t) = \sum_{m:t_m<t} I(i_m=i)$
\item Sender in-degree: $s_{ij2}(t) = \sum_{m:t_m<t} I(j_m=i)$
\item Receiver out-degree: $s_{ij3}(t) = \sum_{m:t_m<t} I(i_m=j)$
\item Receiver in-degree: $s_{ij4}(t) = \sum_{m:t_m<t} I(j_m=j)$
\item Reciprocity (AB-BA): $s_{ij5}(t) = \sum_{m:t_m<t} I(i_m=i,j_m=j,i_{v_{ijm}}=j,j_{v_{ijm}}=i)$
\item Turn-taking (AB-BY): $s_{ij6}(t) = \sum_{m:t_m<t} I(i_m=i,j_m=j,i_{v_{ijm}}=j,j_{v_{ijm}}=i)$
\item Turn* (AB-XA): $s_{ij7}(t) = \sum_{m:t_m<t} I(i_m=i,j_m=j,i_{v_{ijm}}=j,j_{v_{ijm}}=i)$
\item  (AB-XB): $s_{ij8}(t) = \sum_{m:t_m<t} I(i_m=i,j_m=j,i_{v_{ijm}}=j,j_{v_{ijm}}=i)$
\item  (AB-AY): $s_{ij9}(t) = \sum_{m:t_m<t} I(i_m=i,j_m=j,i_{v_{ijm}}=j,j_{v_{ijm}}=i)$
\item TODO: Shared contacters, contactees (also Facebook-like terms.) 
\item TODO: Recency or functions of time-windows
\end{enumerate}

@@

\begin{enumerate}
\item $t_0 = 0$, all $s_{ij} = 1$
\item For $i = 1, \ldots, N$, draw $z_i \sim \mbox{Categorical}(\theta)$
\item Initialize $v = 0$
\item For $m = 1, \ldots, M$:
  \begin{enumerate}
  \item For all $r \ne i$:
    \begin{enumerate}
    \item Compute $s_{rj}^*$ using event $0$ through $v_{rj}$
    \item Set $\lambda_{rj}^* \leftarrow \exp\{ \beta_{z_r,z_j} s_{rj}^*\}$ 
    \end{enumerate}
  \item For all $r \ne j$:
    \begin{enumerate}
    \item Compute $s_{ir}^*$ using event $0$ through $v_{ir}$
    \item Set $\lambda_{ir}^* \leftarrow \exp\{ \beta_{z_i,z_r} s_{ir}^*\}$ 
    \end{enumerate}    
  \item Draw $t_m \sim \mbox{Exponential}\left(\sum_{ij} \lambda_{ij}^*\right)$
  \item Draw $(i_m,j_m) \sim \mbox{Categorical}\left(\lambda_{ij}^* / \sum_{ij}\lambda_{ij}^*\right)$
  \item Save $s_{ij}(t_m) \leftarrow s_{ij}^*$
  \item Save $\lambda_{ij}(t_m) \leftarrow \lambda_{ij}^*$
  \end{enumerate}
\end{enumerate}
